Retrofitting Effect Handlers onto OCaml

### 2021-03-14 retro-concurrency.tex
cite pottier 21 on separation logic for effect handlers
Update code to use intro exceptions

\begin{lstlisting}[numbers=left]
let run main =
  let runq = Queue.create () in
	let suspend k =
	  Queue.push (fun () -> continue k ()) runq
	in
  let pending_reads = ref [] in
  let rec run_next () =
    match Queue.take_opt runq with
    | None ->
      begin match !pending_reads with
      | [] -> () (* no pending reads *)
      | todo ->
        let compl,todo = do_reads todo in
        List.iter (fun (str,k) ->
          Queue.push (fun () -> continue k str) runq) compl;
        pending_reads := todo;
        run_next ()
      end
    | Some f -> f ()
  in
  let rec spawn f =
    match f () with
		...
    | effect (Read ic) k ->
        pending_reads := (ic,k) :: !pending_reads;
        run_next ()
  in
  ...
\end{lstlisting}

@@

Multi-shot continuations are difficult to
reason about in the presence of linear resources such as file descriptors.
Recall that OCaml programs are written with the expectation that function calls
return \emph{exactly-once} either normally or exceptionally. With multi-shot
continuations, they may return \emph{more than once}, leading to errors such as
double-freeing malloced memory. Since our goal is to retrofit effect handlers
onto OCaml, we relegate continuation cloning to the OCaml |Obj| module, which
contains unsafe features.

>>>>

\begin{lstlisting}[numbers=left]
let run main =
  let runq = Queue.create () in
  let suspend k =
    Queue.push (fun () -> continue k ()) runq
  in
  let pending_reads = ref [] in
  let rec run_next () =
    match Queue.pop runq with
    | f -> f ()
    | exception Queue.empty ->
      begin match !pending_reads with
      | [] -> () (* no pending reads *)
      | todo ->
        let compl,todo = do_reads todo in
        List.iter (fun (str,k) ->
          Queue.push (fun () -> continue k str) runq) compl;
        pending_reads := todo;
        run_next ()
      end
  in
  let rec spawn f =
    match f () with
    ...
    | effect (Read ic) k ->
        pending_reads := (ic,k) :: !pending_reads;
        run_next ()
  in
  ...
\end{lstlisting}

@@

Multi-shot continuations are difficult to
reason about in the presence of linear resources such as file descriptors~\cite{de-vilhena-pottier-21}.
Recall that OCaml programs are written with the expectation that function calls
return \emph{exactly-once} either normally or exceptionally. With multi-shot
continuations, they may return \emph{more than once}, leading to errors such as
double-freeing malloced memory. Since our goal is to retrofit effect handlers
onto OCaml, we relegate continuation cloning to the OCaml |Obj| module, which
contains unsafe features.
